---
title: "R Coding Exercise"
---

install.packages("dslabs")
library("dslabs")

help(gapminder)

#Used to examine struture/layout of data
str(gapminder)

#Used to examine data summary
summary(gapminder)

#Used to determine what type of class "gapminder" is 
class(gapminder)

# Downloaded tidyverse and dplyr packages to make it easier to code
install.packages("tidyverse")
library("tidyverse")
install.packages("dplyr")
library("dplyr")


# Generate new objects containing values for African countries only
africadata <- gapminder %>%
  filter(continent == "Africa")
africadata <- africadata %>% filter(continent == "Africa")

# Using str and summary to view structure and summary of new object "africadata"
str(africadata)
summary(africadata)

# Create a new object with life expectancy and infant mortality from africadata
africa_infanmort_lifexpect <- africadata %>%
  select(infant_mortality, life_expectancy)

# View structure and summary of the mortality and life expectancy data
str(africa_infanmort_lifexpect)
summary(africa_infanmort_lifexpect)

# Create a new object with population and life expectancy from African data
africa_pop_lifexpect2 <- africadata %>%
  select(population, life_expectancy)

# View structure and summary of the population and life expectancy data
str(africa_pop_lifexpect2)
summary(africa_pop_lifexpect2)


#plot life expectancy as a function of infant mortality and as a function of population size

library("ggplot2")

# Plot life expectancy as a function of infant mortality
Africa_LE_vs_IM <- ggplot(africa_infanmort_lifexpect, aes(x = infant_mortality, y = life_expectancy)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Life Expectancy vs Infant Mortality",
       x = "Infant Mortality",
       y = "Life Expectancy")
       

# Plot life expectancy vs population with population on a log scale
Africa_LE_vs_Pop_log <- ggplot(africa_pop_lifexpect2, aes(x = population, y = life_expectancy)) +
    geom_point() +
    theme_minimal() +
    labs(title = "Life Expectancy vs Population",
         x = "Population (Log Scale)",
         y = "Life Expectancy")
scale_x_log10()
plot(Africa_LE_vs_Pop_log)

# The streaks you see in the graph are because there appears to be a general trend for countries with larger populations to have higher life expectancies. 

## creating object from africadata for the year 2000
missing_years <- unique(africadata[is.na(africadata$infant_mortality), "year"])
print(missing_years)

```{r}
# Filter data for year == 2000
africadata_2000 <- africadata %>%
  filter(year == "2000")


#Examine the structure and summary of object africa2000
str(africa2000)
summary(africa2000)

```

## new plots using only data from the year 2000 in the africadata file
```{r}
# plot infant mortality vs life expectancy for year=2000
library(ggplot2)

LE_vs_IM_2000 <- ggplot(africa2000, aes(x = infant_mortality, y = life_expectancy)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Life Expectancy vs Infant Mortality, year=2000",
       x = "Infant Mortality",
       y = "Life Expectancy")
plot(LE_vs_IM_2000)

# plot life expectancy vs population (log10) for year=2000
LE_vs_Pop_log_2000 <- ggplot(africa2000, aes(x = population, y = life_expectancy)) +
    geom_point() +
    theme_minimal() +
    labs(title = "Life Expectancy vs Population Year = 2000",
         x = "Population (Log Scale)",
         y = "Life Expectancy") +
    scale_x_log10()

plot(LE_vs_Pop_log_2000)

```

# Using lm() to create fit life expectancy as the outcome, and infant mortality as the predictor
```{r}
fit1 <- lm(life_expectancy ~ infant_mortality, data = africa2000)

summary(fit1)


```
## Using lm() to fit life expectancy as the outcome, and population size as the predictor
```{r}
fit2 <- lm(life_expectancy ~ population, data = africa2000)

summary(fit2)

```
# Interpreting the results for the linear models
# Fit1 demonstrates the relationship between life expectancy and infant mortality using the gapminder database, selecting for African counries in the year 2000. The low p-value of 2.826e-08, suggests there is a statistically significant relationship between life expectancy and infant mortality

# Fit2 demonstrates the relationship between life expectancy and population size using the gapminder database, selecting for African counries in the year 2000. The high p-value of 0.6159, suggests there is not a statistically signicant relationship between life expectancy and population size. 


```{r}

# This section contributed by Clarke Miller.

#Explore the dataset.

#load dslabs package
library("dslabs")
library(dplyr)
library(ggplot2)
library(plotly)


#look at help file for temp_carbon data
help(temp_carbon)

#get an overview of data structure
str(temp_carbon)

#get a summary of data
summary(temp_carbon)

#determine the type of object gapminder is
class(temp_carbon)

View(temp_carbon)

```
```{r}

#Do some processing and cleaning.

narrow_time <- temp_carbon %>% filter(year==c(1880:2018))
narrow_time <- narrow_time %>% filter(year==c(1880:2014))

#get an overview of data structure and get a summary of data
str(narrow_time)
summary(narrow_time)
View(narrow_time)

```
```{r}
#Making a few exploratory figures.

#year and carbon emissions.
carbontime <- narrow_time %>% 
  select(year, carbon_emissions)
  summary(carbontime)
  str(carbontime)
  View(carbontime)
  
#Everybody needs a thrill!
oceantime <- narrow_time %>%
 select(year, ocean_anomaly)
 summary(oceantime)
 str(oceantime)
 View(oceantime)
 


#Making those graphs!
library(ggplot2)

plot_carbon <- ggplot(carbontime, aes(x=year, y=carbon_emissions)) + geom_point() + theme_classic() + labs(title = "Carbon Emmissions from 1880 to 2018", x="Year", y="Carbon Emmissions (Metric tons)")

plot_carbon

plot_ocean <- ggplot(oceantime, aes(x=year, y=ocean_anomaly)) + geom_point() + theme_classic() + labs(title = "Annual mean temperature anomaly over ocean in degrees Celsius relative to the 20th century mean temperature. 1880-2018.", x="Year", y="Temp (Celcius)")

plot_ocean

#Simple Statistical Models
fit3 <- lm(carbon_emissions ~ ocean_anomaly, data=narrow_time)
summary(fit3)

anova_result <- aov(carbon_emissions ~ ocean_anomaly, data=narrow_time)
summary(anova_result)

#It is apparent from the models below that carbon emissions have been increasing since 1880 with a strong tend upwards starting in about 1920. Since about 1920 the ocean temperature has also been increasing.  It is apparent from the following there may be a correlation between rising carbon dioxide release and rising ocean temperatures.  

```


