---
title: "Data-exercise"
---

This exercise is designed to give you practice with exploring text data sets (a _complex data type_). We will load, process, and engage in some descriptive/exploratory analysis of the data.

The data this week comes from the appa R package created by Avery Robbins. The original data came from the Avatar Wiki, and the example code used to scrape this data set is also covered on Avery's blog along with a quick exploration of the available data in a separate blog post.


I'm not sure if I will need these packages but I will load them all just to be sure 
```{r}
library(tidytext)
library(dplyr)
library(tidyverse)
library(stringr)
library(tibble)
library(knitr)
library(here)
library(ggplot2)
library(naniar)
```

First, we will install and load the dataset.
#Note: I did have to installl a package called 'devtools' and also use devtools::install_github("averyrobbins1/appa") in order to use the package "appa"
```{r}
tuesdata <- tidytuesdayR::tt_load('2020-08-11')
tuesdata <- tidytuesdayR::tt_load(2020, week = 33)

avatar <- tuesdata$avatar

library(tidyverse)
library(appa)

avatar <- appa::appa

scene_description <- avatar %>% 
  select(id, scene_description) %>% 
  unnest_longer(scene_description) %>% 
  filter(!is.na(scene_description))

```
Using 'dplyr' to explore the Avatar dataset
```{r}
dplyr::glimpse(avatar)

head(avatar) #This allows me to display the first n rows present in the input data frame.
book_titles <- avatar %>% distinct(book) #There are four books in this series and I would like to know if all books are included in this dataset

print(book_titles)
``` 
The tibble above demonstrates that only three of the books are included in this dataset.


My favorite character in this series is Uncle Iroh (listed as Iroh under characters). I was admittedly obsessed with this character as child ;). 

I would like to see how many times Uncle Iroh is mentioned in this dataset. 
```{r}
characters_found <- grep("Iroh", avatar$character, fixed = TRUE)
print(characters_found)
```
I am able to discern that the numbers in [] relate to the book number. the data correlates the time Iroh is mentioned to the specific book number.  

```{r}

iroh_water<- avatar%>%
  filter(book == "Water" & character == "Iroh")
  
```

I will now attempt to organize this output into a table.
```{r}
Iroh_freq_df <- as.data.frame(iroh_water)
print(Iroh_freq_df)
```
I will now attempt to search for phrases Uncle Iroh says in Water.
```{r}
Iroh_phrases <- grep("Prince Zuko|More tea", iroh_water$character_words, fixed = TRUE)
print(Iroh_phrases)
```
According to the output it is used 0 times. However, both phrases are mentioned. I'm not sure how to correctly format this code as it worked but the response is incorrect. 