---
title: "cdcdata-exercise"
author: "Arlyn Santiago"
editor: visual
---

# CDC Data Exploration: Quarterly Provisional Estimates for Selected Birth Indicators

### Description of the data

This dataset comes from the National Center for Immunization and Respiratory Diseases (NCIRD). The data is a quarterly release of provisional estimates of selected reproductive indicators including general fertility rates, age-specific birth rates, total and low-risk cesarean delivery rates, preterm birth rates, and other gestational age categories. Data on RSV vaccinations among pregnant persons was available starting September 22, 2023, and includes doses received starting September 24, 2023.

The dataset can be found on [CDC website](https://data.cdc.gov/Pregnancy-Vaccination/Weekly-Respiratory-Syncytial-Virus-RSV-Vaccination/g4jn-64pd/about_data.).

### Reading in the data

Here, I loaded all the packages I will use for this data exercise, listed below. I also read in the data from the csv file, and I called the data `cdcdata`. We can see that there are 7 variables and 162 observations. This dataset has multiple indicators stacked on top of each other. For example, within the data there are columns for "Week_Ending_Date", "Race and Ethnicity," etc.,

```{r}
#Loading packages
library(tidyverse) #This includes ggplot2, tidyr, readr, dplyr, stringr, purr, forcats
library(knitr)
library(here)
library(dplyr)
library(tidyr)
library(skimr)
library(ggplot2)
library(kableExtra) #This is used to build common complex tables and manipulate table styles. 

#Reading in the csv file
cdcdata <- read_csv(here("cdcdata-exercise", "Weekly_Respiratory_Syncytial_Virus__RSV__Vaccination_Coverage_among_Pregnant_Persons_by_Race_and_Ethnicity_20240207.csv"))

#Checking the packaging (displaying first and last few rows of data)
nrow(cdcdata)
ncol(cdcdata)

#Showing the structure
str(cdcdata)

#Get column names
colnames(cdcdata)
```

I am seeing many Unknown Values within my dataset in the Race/Ethnicity column so I will try and filter those out with the assistance of AI written code.

```{r}
#Examining data again 
head(cdcdata)

#Filtering Unknown in Column Race and Ethnicity
clean_cdc_data <- cdcdata %>%
  filter(!`Race and Ethnicity` %in% c("Unknown","Overall"))

# Print the filtered data
print(clean_cdc_data)

# Explore clean_cdc_data
head(clean_cdc_data)

```

I will now try to perform some exploratory/descriptive analysis on this cleaned dataset.

```{r}
# Seeing if I can identify overall population of of pregnant persons aged 18-49 years
cdcdata1 <- clean_cdc_data %>%
 group_by(`Race and Ethnicity`) %>%
  reframe(total_denom = sum(Denominator))
print(cdcdata1)

```

```{r}
# Seeing if I can compare the vaccination coverage percentage among races
clean_cdc_data <- clean_cdc_data %>%
  mutate(Numerator = Percent/ 100 * Denominator) # Used this to create a new variable with just the numerator to calculate overall proportion

cdcdata2 <- clean_cdc_data %>%
group_by(`Race and Ethnicity`) %>%
  reframe(total_denom = sum(Denominator),total_numerator = sum(Numerator), overall_percent= total_numerator/total_denom * 100)

print(cdcdata2)
```

```{r}
# Creating a plot of cdcdata2

p1 <- ggplot(cdcdata2, aes(x = `Race and Ethnicity`, y = overall_percent)) + geom_bar(stat = "identity", position = "dodge") +
labs(title = "Distribution of RSV Vaccine Coverage Using Race and Ethnicity", x = "Race and Ethnicity", y = "Percent Vaccinated") +  
guides(x =  guide_axis(angle = 90)) +
theme_minimal()    

p1

save_data_location <- here::here("cdcdata-exercise","cdcdata2.rds")
saveRDS(cdcdata2, file = save_data_location)


```

## This section contributed by Clarke Miller

```{r}

# Call a whole bunch -O- libraries! 
library(here) 
library(dplyr) 
library(ggplot2) 
library(skimr) 
library(gtsummary) 
library(knitr) 
library(tidyr) 
library(broom) 
library(readxl)
library(synthpop)
library(kableExtra)
library(shiny)

#Okay, I caved and used ChatGPT to help figure this part out.  Especially the sythpop stuff. Blah.

# Load your existing data set (replace 'your_dataset.csv' with your actual file path)
existing_data <- read.csv(here("cdcdata-exercise", "Weekly_Respiratory_Syncytial_Virus__RSV__Vaccination_Coverage_among_Pregnant_Persons_by_Race_and_Ethnicity_20240207.csv"))

head(existing_data)

print(existing_data)


#Filtering Unknown in Column Race and Ethnicity
clean_existing_data <- existing_data %>%
  filter(!Race.and.Ethnicity %in% c("Unknown","Overall"))

# Print the filtered data
print(clean_existing_data)

# Explore clean_cdc_data
head(clean_existing_data)

# Mutate!
clean_existing_data <- clean_existing_data %>%
  mutate(Numerator = Percent/ 100 * Denominator) 

cdcdata3 <- clean_existing_data %>%
group_by(Race.and.Ethnicity) %>%
  reframe(total_denom = sum(Denominator),total_numerator = sum(Numerator), overall_percent= total_numerator/total_denom * 100)

head(cdcdata3)

print(cdcdata3)

CDC_plot2 <- ggplot(cdcdata3, aes(x = Race.and.Ethnicity, y = overall_percent)) + geom_bar(stat = "identity", position = "dodge", color = "red") +
labs(title = "Distribution of RSV Vaccine Coverage Using Race and Ethnicity (Original Data)", x = "Race and Ethnicity", y = "Percent Vacinated") +  
guides(x =  guide_axis(angle = 90)) +
theme_minimal()    

CDC_plot2

save_data_location <- here::here("cdcdata-exercise","cdcdata3.rds")
saveRDS(cdcdata3, file = save_data_location)

```

```{r}
# Call a whole bunch -O- libraries! 
library(here) 
library(dplyr) 
library(ggplot2) 
library(skimr) 
library(gtsummary) 
library(knitr) 
library(tidyr) 
library(broom) 
library(readxl)
library(synthpop)
library(kableExtra)
library(shiny)


# Use synthpop to generate synthetic data
synthetic_data <- syn(cdcdata3, method = "cart")

# View the synthetic data
head(synthetic_data)

print(synthetic_data)

save_data_location <- here::here("cdcdata-exercise","synthetic_data.rds")
saveRDS(synthetic_data, file = save_data_location)

CDC_plot5 <- ggplot(synthetic_data, aes(x = Race.and.Ethnicity, y = overall_percent)) + geom_bar(stat = "identity", position = "dodge", color="orange") +
labs(title = "Distribution of RSV Vaccine Coverage Using Race and Ethnicity (Synthetic Data)", x = "Race and Ethnicity", y = "Percent Vaccinated") +  
guides(x =  guide_axis(angle = 90)) +
theme_minimal()  

CDC_plot5

```
